Improved Bounds for Online Balanced Repartitioning

### 2020-04-22 paper/disc20clustering.tex
k=2 complete proof

\begin{theorem}
	The algorithm is 4-competitive.
\end{theorem}
\begin{proof}
\mahmoud{still proof sketch}
We charge both OPT and Alg on each request that triggers a collocation.
Let $C^{t}_{\set{u,v}}$ denote the  value of $C_{\set{u,v}}$ immediately after arrival of $\sigma_t$.
Assume on request $\sigma_{t}=\set{u,v}$,
Alg collocates the pair,
which means $ C^t_{\set{u,v}} = \lambda$ immediately before Alg resets the counter.
Let $t''<t$ be the latest time (i.e.~request) when Alg collocates $\set{u,v}$ before $t$.
Let $t', t'' < t' < t$ be the  time when Alg separates $\set{u,v}$
in order to collocate some other pair sharing a node,
e.g., the pair $\set{u,v'}, v' \neq v$.
%e.g., the request $\sigma_{t'} = \set{u,v'}$.
%By Alg's definition,
%$C^{t'}_{\set{u,v'}}  = C^t_{\set{u,v}} = \lambda$.
We denote the set of all requests to $\set{u,v}$
between times $t'$ and $t$ by $\sigma_{\set{u,v}}[t',t]$.
Thus,
$| \sigma_{\set{u,v}}[t',t] | = \lambda$.
The overall cost of requests that increases the counter,
i.e.~requests in $\sigma_{\set{u,v}}[t',t]$,
and the moving cost incurred by Alg,
 is $\lambda + 2\alpha$.
%That is, the cost of all requests in $\sigma_{\set{u,v}}[t',t]$.
We distinguish three cases for the cost incurred by OPT with regards to $\set{u,v}$.
i)
OPT has $\set{u,v}$ separated through all requests in $\sigma_{\set{u,v}}[t',t]$
and therefore it pays $\lambda$.
ii)  
OPT collocates $\set{u,v}$ some time during $[t'',t]$.
Recall that whenever Alg collocates a pair,
we charge OPT only for requests up to that point (i.e., not for later requests).
Hence,
we charge the cost of  this collocation to OPT for the first time.
In particular,
we charge OPT the cost of moving one node to the cluster of the other,
e.g.~$u$ to $v$, w.l.o.g.
Therefore in this case OPT pays $\alpha$.
iii)
OPT has $\set{u,v}$ collocated before $t''$ and keeps them collocated during $[t'',t]$.
Obviously in this case OPT doesn't pays any cost for requests in $\sigma_{\set{u,v}}[t'',t]$.
However,
OPT must pay $\lambda$ for requests to $\set{u,v'}$.
By induction,
at time $t'$,
we charge OPT $\alpha$ (out of $\lambda=2\alpha$) for this pair.
Therefore,
we can charge OPT another $\alpha$ to account for requests $\sigma_{\set{u,v}}[t'',t]$.
It can be easily verified that any deviation from these cases only increases the cost to OPT.
Thus,
OPT pays at least $\alpha$ for each collocation by Alg
which gives the competitive ratio 
$(\lambda + 2\alpha)/ \alpha = 4$.

>>>>

\begin{theorem} \label{thm:k=2}
	The algorithm is (strictly) 4-competitive.
\end{theorem}

We charge both OPT and Alg whenever Alg collocates a pair (i.e., on a collocation event).
%More precisely,we charge them only the cost of the collocation, per such event.
%Hence,we charge the cost of  this collocation to OPT for the first time.
Alg collocates a pair always with a swap (i.e., two simultaneous migrations),
which  costs $2\alpha$.
%We charge OPT only for requests up to the one that triggers a collocation (i.e., not for later requests on the same pair).
OPT may save some costs by collocating multiple pairs at once (before they incur too much cost),
paying the price of only one migration per pair.
This observation suggests that we can't charge OPT more than $\alpha$ per collocation,
even when OPT actually pays more that that.
E.g.,
OPT can collocate two costly pairs with a single swap if it has both pairs on the cut of same two clusters.
Later, Alg collocates those pairs with two swaps (see Figure \ref{fig:TBD}).
Hence,
whenever OPT collocates a pair,
we charge OPT only the cost of moving a single node to the cluster of the other,
i.e., $\alpha$ (in contrast to $2\alpha$ for Alg).

\begin{proof}[Proof of Theorem \ref{thm:k=2}]
	\mahmoud{still proof sketch}
	Let $C^{t}_{\set{u,v}}$ denote the  value of $C_{\set{u,v}}$ immediately after arrival of $\sigma_t$.
	Assume on request $\sigma_{t}=\set{u,v}$,
	Alg collocates the pair,
	which means $ C^t_{\set{u,v}} = \lambda$ immediately before Alg resets the counter.
	Let $\sigma_{t''}, t''<t$ be the last request (before $\sigma_{t}$)
	on which Alg collocates $\set{u,v}$.
	Let $t'$ s.t.~$t'' < t' < t$ be the  time when Alg separates $\set{u,v}$
	in order to collocate some other pair that shares a node with $\set{u,v}$,
	denoted by $\set{u,v'}, v' \neq v$.
	%e.g., the request $\sigma_{t'} = \set{u,v'}$.
	%By Alg's definition,
	%$C^{t'}_{\set{u,v'}}  = C^t_{\set{u,v}} = \lambda$.
	We denote the set of all requests to $\set{u,v}$
	between times $t'$ and $t$ by $\sigma_{\set{u,v}}[t',t]$.
	Thus,
	$| \sigma_{\set{u,v}}[t',t] | = \lambda$.
	The overall cost of requests that increase the counter,
	i.e.~requests in $\sigma_{\set{u,v}}[t',t]$,
	and the moving cost,
	incurred by Alg is $\lambda + 2\alpha$.
	%That is, the cost of all requests in $\sigma_{\set{u,v}}[t',t]$.
	We distinguish three cases for the cost incurred by OPT due to $\set{u,v}$.
	Case i)
	OPT has $\set{u,v}$ separated (in OPT's configuration) during $[t',t]$
	and therefore it pays $\lambda$.
	We charge OPT $\lambda/2 = \alpha$.
	The remaining cost will be used in a latter case.
	Case ii)  
	OPT collocates $\set{u,v}$ some time during $[t'',t]$.
	Any cost that we charge to OPT upon $\sigma_{t''}$,
	accounts for OPT's actions before $t''$.
	Therefore in this case,
	we charge OPT the (unaccounted) cost $\alpha$.
	Case iii)
	OPT has $\set{u,v}$ collocated before $t''$ and keeps them collocated during $[t'',t]$.
	Obviously in this case OPT doesn't pays any cost for requests in $\sigma_{\set{u,v}}[t'',t]$ and
	we can't charge it directly.
	However,
	OPT must pay $\lambda$ for all requests in $\sigma_{\set{u,v'}}[t'',t']$,
	and we charge it partially to OPT at $t'$ (as in Case i).
	Since we always charge OPT no more than $\alpha$,
	we can charge OPT another $\alpha$,
	the remaining cost of requests $\sigma_{\set{u,v'}}[t'',t']$.
	Intuitively,
	at $t$,
	we charge OPT $\alpha$ for separating $\set{u,v}$ at $t'$.
	It can be easily verified that any deviation from these cases only increases the cost to OPT.
	Thus,
	OPT pays at least $\alpha$ for each collocation performed by Alg,
	which gives the competitive ratio 
	$(\lambda + 2\alpha)/ \alpha = 4$.

	It remains to charge both Alg and OPT for requests that do not lead to collocations when the sequence ends.
	Consider  pairs $\set{u,v}$ and $\set{u,v'}$ and
	times $t'' < t' < t = t_m$ as defined previously.
	Assume $\sigma_{\set{u,v}}[t'', t_m] \leq \lambda - 1$.
	This means Alg has the pair separated at $t_m$
	and paid the (communication) cost $r_{\set{u,v}} := | \sigma_{\set{u,v}}[t'', t_m]|$.
	If OPT has the pair separated during $[t'',t_m]$ then it pays $r_{\set{u,v}}$.
	If OPT collocates the pair some time during $[t'', t]$ then analogous to Case (ii),
	we charge it $\alpha$.
	Otherwise,
	if OPT collocates the pair before $t''$,
	then we charge it $\alpha$ analogously to Case (iii).
	Hence, 
	$Cost(OPT)_{\set{u,v}} / Cost(Alg)_{\set{u,v}} = r_{\set{u,v}}/ \alpha \leq (\lambda-1) / \alpha < 2$.
	The claim follows from here.
\end{proof}
